---
title: "506 PS2 Yeomans"
author: "Sydney Yeomans"
format:
  html:
    embed-resources: true
editor: visual
---

## Problem 1

Consider a 1-dimensional random walk with the following rules:

1. Start at 0.
2. At each step, move $+1$ or $-1$ with $50$/$50$ probability.
3. If $+1$ is chosen, $5\%$ of the time move $+10$ instead.
4. If $-1$ is chosen, $20\%$ of the time move $-3$ instead.
5. Repeat steps $2$ - $4$ $n$ times.

Write a function to determine the end position of this random walk.

```{r}
#| echo: true

```

a. Implement the random walk in these three versions: 1. using a loop, 2. using built-in R vectorized functions. (Using no loops.), 3. Implement the random walk using one of the `apply` functions. 

```{r}
#| echo: true

```

b. Demonstrate that the three versions can give the same result. Show this for both $n=10$ and $n=100$. (You will need to add a way to control the randomization.)

```{r}
#| echo: true

```

c. Use the `microbenchmark` package to clearly demonstrate the speed of the implementations. Compare performance with a low input (1,000) and a large input (100,000). Discuss the results.

```{r}
#| echo: true

```

d. What is the probability that the random walk ends at 0 if the number of steps is $10$? $100$? $1000$? Defend your answers with evidence based upon a Monte Carlo simulation.

```{r}
#| echo: true

```


## Problem 2

The number of cars passing an intersection is a classic example of a Poisson distribution. At a particular intersection, Poisson is an appropriate distribution most of the time, but during rush hours (hours of 8am and 5pm) the distribution is really normally distributed with a much higher mean.

Using a Monte Carlo simulation, estimate the average number of cars that pass an intersection under the following assumptions:

1. From midnight until 7 AM, the distribution of cars is Poisson with mean $1$.
2. From 9am to 4pm, the distribution of cars is Poisson with mean $8$.
3. From 6pm to 11pm, the distribution of cars is Poisson with mean $12$.
4. During rush hours (8am and 5pm), the distribution of cars is Normal with mean $60$ and variance $12$.

Accomplish this without using any loops.


```{r}
#| echo: true

```


## Problem 3

Use the following code to download the YouTube Superbowl commercials data:

```{r}
#| echo: true
youtube <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')
```

The research question for this project is to decide which of several attributes, if any, is associated with increased YouTube engagement metrics.

a. Remove any column that might uniquely identify a commercial. This includes but isn’t limited to things like brand, any URLs, the YouTube channel, or when it was published.

Report the dimensions of the data after removing these columns.

```{r}
#| echo: true

```

b. For each of the following variables, examine their distribution. Determine whether i) The variable could be used as is as the outcome in a linear regression model, ii) The variable can use a transformation prior to being used as the outcome in a linear regression model, or iii) The variable would not be appropriate to use as the outcome in a linear regression model.

For each variable (view counts, like counts, dislike counts, favorite counts, comment counts), report which category it falls in. If it requires a transformation, carry such a transformation out and use that transformation going forward.

```{r}
#| echo: true

```

c. For each variable in part b. that are appropriate, fit a linear regression model predicting them based upon each of the seven binary flags for characteristics of the ads, such as whether it is funny. Control for year as a continuous covariate.

Discuss the results. Identify the direction of any statistically significant results.

```{r}
#| echo: true

```

d. Consider only the outcome of view counts. Calculate $\hat{\beta}_1$ manually (without using `lm`) by first creating a proper design matrix, then using matrix algebra to estimate $\beta$. Confirm that you get the same result as `lm` did in part c.

```{r}
#| echo: true

```




