---
title: "PS2 506 Yeomans"
author: "Sydney Yeomans"
format:
  html:
    embed-resources: true
editor: visual
---

## Problem 1

Consider a 1-dimensional random walk with the following rules:

1. Start at 0.
2. At each step, move $+1$ or $-1$ with $50$/$50$ probability.
3. If $+1$ is chosen, $5\%$ of the time move $+10$ instead.
4. If $-1$ is chosen, $20\%$ of the time move $-3$ instead.
5. Repeat steps $2$ - $4$, $n$ times.

Write a function to determine the end position of this random walk.

```{r}
#| echo: true
#' Function that creates 1-dimensional random walk rules listed above.
#'
#' @param : The number of steps
#' @return : The final position of the walk
random_walk <- function(n) {
  initial_step <- 0 
  for (i in 1:n) {}
}
```

a. Implement the random walk in these three versions: 1. using a loop, 2. using built-in R vectorized functions. (Using no loops.), 3. Implement the random walk using one of the `apply` functions. 

```{r}
#| echo: true

```

b. Demonstrate that the three versions can give the same result. Show this for both $n=10$ and $n=100$. (You will need to add a way to control the randomization.)

```{r}
#| echo: true

```

c. Use the `microbenchmark` package to clearly demonstrate the speed of the implementations. Compare performance with a low input (1,000) and a large input (100,000). Discuss the results.

```{r}
install.packages("microbenchmark")
```

```{r}
#| echo: true
library("microbenchmark")

```

d. What is the probability that the random walk ends at 0 if the number of steps is $10$? $100$? $1000$? Defend your answers with evidence based upon a Monte Carlo simulation.

```{r}
#| echo: true

```


## Problem 2

The number of cars passing an intersection is a classic example of a Poisson distribution. At a particular intersection, Poisson is an appropriate distribution most of the time, but during rush hours (hours of 8am and 5pm) the distribution is really normally distributed with a much higher mean.

Using a Monte Carlo simulation, estimate the average number of cars that pass an intersection under the following assumptions:

1. From midnight until 7 AM, the distribution of cars is Poisson with mean $1$.
2. From 9am to 4pm, the distribution of cars is Poisson with mean $8$.
3. From 6pm to 11pm, the distribution of cars is Poisson with mean $12$.
4. During rush hours (8am and 5pm), the distribution of cars is Normal with mean $60$ and variance $12$.

Accomplish this without using any loops.

```{r}
#| echo: true
library(stats)
set.seed(23)
#rpois(n, lambda = 1): 00:00 - 07:00
#rnorm(n, mean = 60, sd = 12): 08:00 - 17:00
#rpois(n, lambda = 8): 09:00 - 16:00
#rpois(n, lambda = 12): 18:00 - 23:00
n <- 10000
midnight_to_seven <- rpois(n, lambda = 1)
rush_hour <- rnorm(n, mean = 60, sd = 12)
nine_to_four <- rpois(n, lambda = 8)
six_to_eleven <- rpois(n, lambda = 12)

total_car_pass <- midnight_to_seven + rush_hour + nine_to_four + six_to_eleven
mean(total_car_pass)
```


## Problem 3

Use the following code to download the YouTube Superbowl commercials data:

```{r}
#| echo: true
youtube <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')
```

The research question for this project is to decide which of several attributes, if any, is associated with increased YouTube engagement metrics.

a. Remove any column that might uniquely identify a commercial. This includes but isn’t limited to things like brand, any URLs, the YouTube channel, or when it was published.

Report the dimensions of the data after removing these columns.

```{r}
#| echo: true
dim(youtube)

#remove any columns that give away what it is
youtube <- subset(youtube, 
            select = -c(brand, 
                        superbowl_ads_dot_com_url, 
                        youtube_url, 
                        id, 
                        etag, 
                        published_at, 
                        title, 
                        description, 
                        channel_title))

dim(youtube)
```

I got rid of these variables because I think they are the only ones that could be used to figure out what the commercial is.

b. For each of the following variables, examine their distribution. Determine whether i) The variable could be used as is as the outcome in a linear regression model, ii) The variable can use a transformation prior to being used as the outcome in a linear regression model, or iii) The variable would not be appropriate to use as the outcome in a linear regression model.

For each variable (view counts, like counts, dislike counts, favorite counts, comment counts), report which category it falls in. If it requires a transformation, carry such a transformation out and use that transformation going forward.

In order to see if these variables are appropriate to use in a linear regression model we want to look at the distribution or density of values in that variable and see if it looks roughly normal, which comes from the fact that ... I will be using histograms because I think they are easier to make and deal with.

```{r}
#| echo: true
hist(youtube$view_count)

```

```{r}
#| echo: true
hist(youtube$like_count)
```
```{r}
#| echo: true
hist(youtube$dislike_count)

```

```{r}
#| echo: true
hist(youtube$favorite_count)
```

```{r}
#| echo: true
hist(youtube$comment_count)
```
From the histograms we can see this distribution for `{r} view_count`, `{r} like_count`, `{r} dislike_count`, and `{r} comment_count` are all heavily right skewed with some potential outliers at very high values along the x-axis. Since these are heavily skewed we should use a log transformation to deal with the extreme values. After this log transformation we should hopefully see something that looks a bit better and more Normal. The histogram for the `{r} favorite_count` is the only odd one out from this group of variables. Looking in the data set we actually see that the only value for favorite count is $0$ or NA, so this one won't be helpful for linear regression. All the other variables should work after a transformation.

```{r}
#| echo: true
#log transformation
youtube$view_count <- log(youtube$view_count)
youtube$like_count <- log(youtube$like_count)
youtube$dislike_count <- log(youtube$dislike_count)
youtube$comment_count <- log(youtube$comment_count)
```

```{r}
#check histogram to see if the transformation worked
hist(youtube$view_count)

```

The histogram with the log transformation looks very symmetric and bell-shaped so we should now be able to use these variables for linear regression.


c. For each variable in part b. that are appropriate, fit a linear regression model predicting them based upon each of the seven binary flags for characteristics of the ads, such as whether it is funny. Control for year as a continuous co-variate.

Discuss the results. Identify the direction of any statistically significant results.

```{r}
#| echo: true
#view counts, like counts, dislike counts, favorite counts, comment counts
view_count_lm <- lm(view_count ~ funny + show_product_quickly 
                      + patriotic + celebrity + danger + animals 
                      + use_sex + year, data = youtube)
like_count_lm <- lm(like_count ~ funny + show_product_quickly 
                      + patriotic + celebrity + danger + animals 
                      + use_sex + year, data = youtube)
dislike_count_lm <- lm(dislike_count ~ funny + show_product_quickly 
                      + patriotic + celebrity + danger + animals 
                      + use_sex + year, data = youtube)
comment_count_lm <- lm(comment_count ~ funny + show_product_quickly 
                      + patriotic + celebrity + danger + animals 
                      + use_sex + year, data = youtube)
```

d. Consider only the outcome of view counts. Calculate $\hat{\beta}_1$ manually (without using `lm`) by first creating a proper design matrix, then using matrix algebra to estimate $\beta$. Confirm that you get the same result as `lm` did in part c.

```{r}
#| echo: true

```

